An Analysis of Performance Bottlenecks in MRI Pre-Processing

### 2024-05-09 paper/paper.tex
updates in abstract

\begin{abstract}
	Magnetic Resonance Image (MRI) pre-processing is a critical step for neuroimaging analysis. However, the computational cost of MRI pre-processing pipelines is a major bottleneck for large cohort studies and some clinical applications. While High-Performance Computing (HPC) and, more recently, Deep Learning have been adopted to accelerate the computations, these techniques require costly hardware and are not accessible to all researchers. Therefore, it is critical to understand the performance bottlenecks of MRI pre-processing pipelines to improve their performance. Using Intel VTune profiler, we characterize the bottlenecks of several commonly used MRI-preprocessing pipelines from the ANTs, FSL, and FreeSurfer toolboxes. We found that few functions contribute to most of the CPU time, and that linear interpolation is the largest contributor. Data access is also  a substantial bottleneck. Last, we identified a bug in the ITK library that impact the performance of ANTs pipeline in single-precision and a potential issue with the OpenMP scaling in FreeSurfer recon-all. Our results provide a reference for future efforts to optimize MRI pre-processing pipelines.
\end{abstract}

@@

\section{Results}
In this section we present the aggregated profiling data obtained with VTune. First, we present a performance overview of the pipelines. We then analysis specific hotspots in the pipelines. Unless specified explicitly, we only present the profile data from using 1-thread since the results from 32-threads are similar.

@@

\subsection{VTune results finalization}
\HL{VTune profiler offered a rich and extensive amount of data for our analysis, which would have been difficult to retrieve otherwise. Thanks to it's low performance overhead, we were able to profile the pipeline with minimal impact on the runtime, allowing the use of more subjects to obtain a more accurate performance profile. The documentation provided by VTune is extensive and well-documented, which helped us to understand the different metrics and events available.}

>>>>

\begin{abstract}
	Magnetic Resonance Image (MRI) pre-processing is a critical step for neuroimaging analysis. However, the computational cost of MRI pre-processing pipelines is a major bottleneck for large cohort studies and some clinical applications. While High-Performance Computing (HPC) and, more recently, Deep Learning have been adopted to accelerate the computations, these techniques require costly hardware and are not accessible to all researchers. Therefore, it is important to understand the performance bottlenecks of MRI pre-processing pipelines to improve their performance. Using Intel VTune profiler, we characterized the bottlenecks of several commonly used MRI-preprocessing pipelines from the ANTs, FSL, and FreeSurfer toolboxes. We found that few functions contributed to most of the CPU time, and that linear interpolation was the largest contributor. Data access was also  a substantial bottleneck. We identified a bug in the ITK library that impacts the performance of ANTs pipeline in single-precision and a potential issue with the OpenMP scaling in FreeSurfer recon-all. Our results provide a reference for future efforts to optimize MRI pre-processing pipelines.
\end{abstract}

@@

\section{Results \& Discussion}
In this section we present the aggregated profiling data obtained with VTune. First, we present a performance overview of the pipelines. We then analysis specific hotspots in the pipelines. Unless specified explicitly, we only present the profile data from using 1-thread since the results from 32-threads are similar.

@@

\subsection{Experiencing using VTune}
\HL{VTune profiler offered a rich and extensive amount of data for our analysis, which would have been difficult to retrieve otherwise. Thanks to it's low performance overhead, we were able to profile the pipeline with minimal impact on the runtime, allowing the use of more subjects to obtain a more accurate performance profile. The documentation provided by VTune is extensive and well-documented, which helped us to understand the different metrics and events available.}
